---
title: "PERMANOVA_plasticity"
author: "Adam_Koziol"
date: '2022-07-22'
output:
  html_document: default
  pdf_document: default
---

```{r include=FALSE}
setwd('~/Documents/GitHub/Metgen_plasticityanalysis/PERMANOVA/')
getwd()
library(tidyverse)
library(vegan)
library(kableExtra)
library(hillR)
library(ggvegan)
library(wesanderson)
library(ggpubr)
library(caret)
library(ape)
library(phytools)
```
Kable execution of PERMANOVA tables
```{r}
table_func <- function(input, str_removal, table_title){
maineffects <- input[c(1,2,3),]
interaction <- input[c(4,5),]
maineffectsrow <- rbind(maineffects, interaction) %>%
  rownames_to_column('Main_effects') %>%
  mutate(Main_effects = str_remove_all(Main_effects, '[$]+')) %>%
  mutate(Main_effects = str_remove_all(Main_effects, str_removal))

maineffects_table <- kable(maineffectsrow, format = 'html', digits = 3, caption = table_title) %>%
  kable_styling(full_width = F)

return(maineffects_table)
}

```

Loading in the data files
```{r include=FALSE}
metadata <- read.csv('../data/metadata.csv', row.names = 1, header = T)

CR_met <- metadata %>%
  filter(species == 'CR') %>%
  column_to_rownames('sample_id') %>%
  mutate(sample_id = rownames(.))

AS_met <- metadata %>%
  filter(species == 'AS') %>%
  column_to_rownames('sample_id') %>%
  mutate(sample_id = rownames(.))

CR_counts <- read.csv('../data/TPM_CR.csv', row.names = 1, header = T)
AS_counts <- read.csv('../data/AS_TPM.csv', row.names = 1, header = T) %>%
  .[,rownames(AS_met)]

CR_tree <- read.tree(file='../data/gtdbtk.bac120.classify_CR.tree') %>%
  keep.tip(row.names(CR_counts)) %>%
  force.ultrametric(CR_tree,method="extend")

AS_tree <- read.tree(file='../data/gtdbtk.bac120.classify_AS.tree') %>%
  keep.tip(row.names(AS_counts)) %>%
  force.ultrametric(AS_tree,method="extend")

AS_products <- read.csv('../damma/outputsR/distilled_table_functions_AS.csv', row.names = 1, header = T) %>%
  .[rownames(AS_counts),]

CR_products <- read.csv('../damma/outputsR/distilled_nonbinary_compounds_CR.csv', row.names = 1, header = T) %>%
  .[rownames(CR_counts),]
```

Taxonomic dissimilarity matrix
```{r include=FALSE}
CR_dissim <- hill_taxa_parti_pairwise(t(CR_counts), q = 1, output = 'matrix') %>% 
  .[[5]] %>%
  as.data.frame %>% 
  mutate_all(~1 - .) %>%
  t() %>% 
  as.dist()

AS_dissim <- hill_taxa_parti_pairwise(t(AS_counts), q = 1, output = 'matrix') %>% 
  .[[5]] %>%
  as.data.frame %>% 
  mutate_all(~1 - .) %>%
  t() %>% 
  as.dist()
```

Phylogenetic dissimilarity matrix
```{r include=FALSE}
CR_dissim_pd <- hill_phylo_parti_pairwise(t(CR_counts), tree = CR_tree, q = 1, output = 'matrix') %>% 
  .[[5]] %>%
  as.data.frame %>% 
  mutate_all(~1 - .) %>%
  t() %>% 
  as.dist()

AS_dissim_pd <- hill_phylo_parti_pairwise(t(AS_counts), q = 1, tree=AS_tree, output = 'matrix') %>% 
  .[[5]] %>%
  as.data.frame %>% 
  mutate_all(~1 - .) %>%
  t() %>% 
  as.dist()
```

Functional dissimilarity matrix
```{r include=FALSE}
CR_products <- CR_products[rownames(CR_counts),]

CR_dissim_fd <- hill_func_parti_pairwise(comm = t(CR_counts),
                                         traits = CR_products,
                                         output = 'matrix',
                                         q=1) %>% 
  .[[5]] %>%
  as.data.frame %>% 
  mutate_all(~1 - .) %>%
  t() %>% 
  as.dist

AS_dissim_fd <- hill_func_parti_pairwise(t(AS_counts), AS_products, output = 'matrix', q=1) %>% 
  .[[5]] %>% 
  as.data.frame %>% 
  mutate_all(~1 - .) %>%
  t() %>% 
  as.dist
```

Perform the PERMANOVA 
```{r echo=TRUE}
###Re-order the metdata
CR_met <- CR_met[colnames(CR_counts),]

CR_perm <- adonis2(CR_dissim ~ CR_met$treatment, strata = CR_met$individual_id, permutations = 9999)

table <- table_func(input = CR_perm,
                    str_removal = 'CR_met',
                    table_title = 'PERMANOVA_CR1_Neutral')

table
```

```{r include=FALSE}
AS_met <- AS_met[colnames(AS_counts),]

AS_perm <- adonis2(AS_dissim ~ AS_met$treatment, strata = AS_met$individual_id, permutations = 9999)

table <- table_func(input = AS_perm,
                    str_removal = 'AS_met',
                    table_title = 'PERMANOVA_AS1_Neutral')

table
```
Perform the phylogenetic PERMANOVA 
```{r echo=TRUE}
CR_perm_phy <- adonis2(CR_dissim_pd ~ CR_met$treatment, strata = CR_met$individual_id, permutations = 9999)

table <- table_func(input = CR_perm_phy,
                    str_removal = 'CR_met',
                    table_title = 'PERMANOVA_CR1_Phylogenetic')

table
```

```{r echo=TRUE}
AS_perm_phy <- adonis2(AS_dissim_pd ~ AS_met$treatment, strata = AS_met$individual_id, permutations = 9999)

table <- table_func(input = AS_perm_phy,
                    str_removal = 'AS_met',
                    table_title = 'PERMANOVA_AS1_Phylogenetic')

table
```

Perform the functional PERMANOVA 
```{r echo=TRUE}
CR_perm_fd <- adonis2(CR_dissim_fd ~ CR_met$treatment, strata = CR_met$individual_id, permutations = 9999)

table <- table_func(input = CR_perm_fd,
                    str_removal = 'CR_met',
                    table_title = 'PERMANOVA_CR1_Functional')

table
```

```{r echo=TRUE}
AS_perm_fd <- adonis2(AS_dissim_fd ~ AS_met$treatment, strata = AS_met$individual_id, permutations = 9999)

table <- table_func(input = AS_perm_fd,
                    str_removal = 'AS_met',
                    table_title = 'PERMANOVA_AS1_Functional')

table
```

######CAP plots - Apodemus neutral CAP
```{r echo=TRUE}
cca_ord <- capscale(formula = AS_dissim ~ AS_met$treatment)

CAP_df <- as.data.frame(vegan::scores(cca_ord, display = "sites")) %>%
  rownames_to_column('sample_id') %>%
  left_join(AS_met, by = 'sample_id') %>%
  column_to_rownames('sample_id')

AS_plot <- ggplot(CAP_df, aes(x = CAP1, y = CAP2, colour = treatment, fill = treatment)) +
  geom_point() +
  scale_colour_manual(values=wes_palette(n=6, name="IsleofDogs1")) +
  scale_fill_manual(values=wes_palette(n=6, name="IsleofDogs1")) +
  theme_classic() +
  theme(axis.title = element_text(face = 'bold',
                             size = 10),
        axis.title.x = element_blank(),
        title = element_text(face = 'italic',
                             size = 14),
        legend.title = element_text(face = 'plain')) +
  geom_hline(yintercept = 0, linetype="dotted") + 
  geom_vline(xintercept = 0, linetype="dotted") +
  ggtitle("A. sylvaticus") +
  labs(fill='Disturbance',
       colour='Disturbance') +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Neutral\n\nCAP2") + 
  xlab("CAP1") +
  stat_ellipse(geom = 'polygon', aes(fill=treatment, colour=treatment), alpha = 0.25)
AS_plot
```
Crocidura neutral CAP
```{r echo=TRUE}
cca_ord <- capscale(formula = CR_dissim ~ CR_met$treatment)

CAP_df <- as.data.frame(vegan::scores(cca_ord, display = "sites")) %>%
  rownames_to_column('sample_id') %>%
  left_join(CR_met, by = 'sample_id') %>%
  column_to_rownames('sample_id')

CR_plot <- ggplot(CAP_df, aes(x = CAP1, y = CAP2, colour = treatment, fill=treatment)) +
  geom_point() +
  scale_colour_manual(values=wes_palette(n=6, name="IsleofDogs1")) +
  scale_fill_manual(values=wes_palette(n=6, name="IsleofDogs1")) +
  theme_classic() +
  theme(axis.title = element_text(face = 'bold',
                             size = 10),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        title = element_text(face = 'italic',
                             size = 14),
        legend.title = element_text(face = 'plain')) +
  geom_hline(yintercept = 0, linetype="dotted") + 
  geom_vline(xintercept = 0, linetype="dotted") +
  labs(title="C. russula") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Neutral\n\nCAP2") + 
  xlab("CAP1") +
  stat_ellipse(geom = 'polygon', aes(fill=treatment, colour=treatment), alpha = 0.25)

CR_plot
```
Apodemus phylogenetic CAP
```{r echo=TRUE}
cca_ord <- capscale(formula = AS_dissim_pd ~ AS_met$treatment)

CAP_df <- as.data.frame(vegan::scores(cca_ord, display = "sites")) %>%
  rownames_to_column('sample_id') %>%
  left_join(AS_met, by = 'sample_id') %>%
  column_to_rownames('sample_id')

AS_plot_pd <- ggplot(CAP_df, aes(x = CAP1, y = CAP2, colour = treatment, fill = treatment)) +
  geom_point() +
  scale_colour_manual(values=wes_palette(n=6, name="IsleofDogs1")) +
  scale_fill_manual(values=wes_palette(n=6, name="IsleofDogs1")) +
  theme_classic() +
  theme(axis.title = element_text(face = 'bold',
                             size = 10),
        axis.title.x=element_blank(),
        title = element_text(face = 'italic',
                             size = 14),
        legend.title = element_text(face = 'plain')) +
  geom_hline(yintercept = 0, linetype="dotted") + 
  geom_vline(xintercept = 0, linetype="dotted") +
  #ggtitle("A. sylvaticus") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Phylogenetic\n\nCAP2") + 
  xlab("CAP1") +
  stat_ellipse(geom = 'polygon', aes(fill=treatment, colour=treatment), alpha = 0.25)

AS_plot_pd
```

Crocidura phylogenetic CAP
```{r echo=TRUE}
cca_ord <- capscale(formula = CR_dissim_pd ~ CR_met$treatment)

CAP_df <- as.data.frame(vegan::scores(cca_ord, display = "sites")) %>%
  rownames_to_column('sample_id') %>%
  left_join(CR_met, by = 'sample_id') %>%
  column_to_rownames('sample_id')

CR_plot_pd <- ggplot(CAP_df, aes(x = CAP1, y = CAP2, colour = treatment, fill = treatment)) +
  geom_point() +
  scale_colour_manual(values=wes_palette(n=6, name="IsleofDogs1")) +
  scale_fill_manual(values=wes_palette(n=6, name="IsleofDogs1")) +
  theme_classic() +
  theme(axis.title = element_text(face = 'bold',
                             size = 10),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        title = element_text(face = 'italic',
                             size = 14),
        legend.title = element_text(face = 'plain')) +
  geom_hline(yintercept = 0, linetype="dotted") + 
  geom_vline(xintercept = 0, linetype="dotted") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("") + 
  xlab("CAP1") +
  stat_ellipse(geom = 'polygon', aes(fill=treatment, colour=treatment), alpha = 0.25)

CR_plot_pd
```

Apodemus functional CAP
```{r echo=TRUE}
cca_ord <- capscale(formula = AS_dissim_fd ~ AS_met$treatment)

CAP_df <- as.data.frame(vegan::scores(cca_ord, display = "sites")) %>%
  rownames_to_column('sample_id') %>%
  left_join(AS_met, by = 'sample_id') %>%
  column_to_rownames('sample_id')

AS_plot_fd <- ggplot(CAP_df, aes(x = CAP1, y = CAP2, colour = treatment, fill = treatment)) +
  geom_point() +
  scale_colour_manual(values=wes_palette(n=6, name="IsleofDogs1")) +
  scale_fill_manual(values=wes_palette(n=6, name="IsleofDogs1")) +
  theme_classic() +
  theme(axis.title = element_text(face = 'bold',
                             size = 10),
        title = element_text(face = 'italic',
                             size = 14),
        legend.title = element_text(face = 'plain')) +
  geom_hline(yintercept = 0, linetype="dotted") + 
  geom_vline(xintercept = 0, linetype="dotted") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Functional\n\nCAP2") + 
  xlab("CAP1") +
  stat_ellipse(geom = 'polygon', aes(fill=treatment, colour=treatment), alpha = 0.25)

AS_plot_fd
```

Crocidura Functional CAP
```{r echo=TRUE}
cca_ord <- capscale(formula = CR_dissim_fd ~ CR_met$treatment)

CAP_df <- as.data.frame(vegan::scores(cca_ord, display = "sites")) %>%
  rownames_to_column('sample_id') %>%
  left_join(CR_met, by = 'sample_id') %>%
  column_to_rownames('sample_id')

CR_plot_fd <- ggplot(CAP_df, aes(x = CAP1, y = CAP2, colour = treatment, fill = treatment)) +
  geom_point() +
  scale_colour_manual(values=wes_palette(n=6, name="IsleofDogs1")) +
  scale_fill_manual(values=wes_palette(n=6, name="IsleofDogs1")) +
  theme_classic() +
  theme(axis.title = element_text(face = 'bold',
                             size = 10),
        axis.title.y = element_blank(),
        title = element_text(face = 'italic',
                             size = 14),
        legend.title = element_text(face = 'plain')) +
  geom_hline(yintercept = 0, linetype="dotted") + 
  geom_vline(xintercept = 0, linetype="dotted") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("CAP2") + 
  xlab("CAP1") +
  stat_ellipse(geom = 'polygon', aes(fill=treatment, colour=treatment), alpha = 0.25)

CR_plot_fd
```

Combine all CAP plots 
```{r eval=FALSE, include=FALSE}
my_combined_CAP <- ggarrange(AS_plot, CR_plot, AS_plot_pd, CR_plot_pd, AS_plot_fd, CR_plot_fd, labels = c('A', 'B', 'C', 'D', 'E', 'F'),
                             ncol=2, nrow = 3, common.legend = T, legend = 'bottom', label.x = -0.018) + bgcolor("White") + border(color = "White", size = 0.8, linetype = NULL)

my_combined_CAP

ggsave('./figures/CAP_full.png', my_combined_CAP, width = 10, height = 7)
ggsave('./figures/CAP_full.pdf', my_combined_CAP, width = 10, height = 7)
```









