---
title: "hill_numbers"
author: "Adam_Koziol"
date: "2022-11-03"
output: html_document
---

Setting the work environment
```{r include=FALSE}
setwd('~/Documents/GitHub/Metgen_plasticityanalysis/hill_numbers/')
library(tidyverse)
library(vegan)
library(nlme)
library(hilldiv)
library(ape)
library(RColorBrewer)
library(ape)
library(cowplot)
library(viridis) 
library(grid)
library(hillR)
library(ggpubr)
library(wesanderson)
library(phytools)
```

Loading metadata and GTDB taxonomy files
```{r}
source('./neatness_functions.R') ###Load in auxillary functions for this data analysis
metadata <- read.csv('../data/metadata.csv', header = T, row.names = 1)
taxa <- read.csv('../data/taxa.csv', sep=',', header=T, row.names=1) 
```

Load Apodemus and Crocidura normalised counts and taxonomic data
```{r}
CR_TPM <- read.csv('../data/TPM_CR.csv')

CR_tree <- read.tree(file='../data/gtdbtk.bac120.classify_CR.tree') %>%
  keep.tip(CR_TPM$X) %>%
  force.ultrametric(CR_tree,method="extend")

AS_TPM <- read.csv('../data/AS_TPM.csv') %>%
  dplyr::select(!ends_with('FAST'))

AS_tree <- read.tree(file='../data/gtdbtk.bac120.classify_AS.tree') %>%
  keep.tip(AS_TPM$X) %>%
  force.ultrametric(AS_tree,method="extend")
```

Create the partitions for overall diversity analyses
```{r}
#####################Apodemus sylvaticus
metadata_ord_AS <- which_ind_2keep(metadata, '^M')
metadata_ord_CR <- which_ind_2keep(metadata, '^S')

apodemus_tpm <- AS_TPM %>%
  column_to_rownames('X')

crocidura_tpm <- CR_TPM %>%
  column_to_rownames('X')

TPM_match_AS <- apodemus_tpm[,colnames(apodemus_tpm) %in% metadata_ord_AS$sample_id]
TPM_match_CR <- crocidura_tpm[,colnames(crocidura_tpm) %in% metadata_ord_CR$sample_id]
mouse_list <- as.character(unique(metadata_ord_AS$individual_id))
shrew_list <- as.character(unique(metadata_ord_CR$individual_id))
```

Overall taxonomic diversity
```{r}
Total_beta_AS_td <- Create_dissims(input_df = TPM_match_AS,
                                   input_metadata = metadata_ord_AS,
                                   input_list = mouse_list,
                                   stored_data = Total_beta_AS_td,
                                   hillR_func = hill_taxa_parti,
                                   extra_input = F)

Total_beta_CR_td <- Create_dissims(input_df = TPM_match_CR,
                                   input_metadata = metadata_ord_CR,
                                   input_list = shrew_list,
                                   stored_data = Total_beta_CR_td,
                                   hillR_func = hill_taxa_parti,
                                   extra_input = F)

Total_beta_td <- rbind(Total_beta_AS_td, Total_beta_CR_td) %>%
  as_tibble

Total_beta_td <- make_overall_csv(Total_beta_td)
write.csv(Total_beta_td, './outputsR/total_betadiversity_td.csv')

Overall_beta <- plot_overall_betas(Total_beta_td)
Overall_beta
ggsave('./figures/individual_betadiversity_td.png', width = 10, height = 7)
ggsave('./figures/individual_betadiversity_td.pdf', width = 10, height = 7)
```

Overall_phylogenetic diversity
```{r}
Total_beta_AS_pd <- Create_dissims(input_df = TPM_match_AS,
                                   input_metadata = metadata_ord_AS,
                                   input_list = mouse_list,
                                   stored_data = Total_beta_AS_pd,
                                   hillR_func = hill_phylo_parti,
                                   extra_input = T,
                                   extra_file = AS_tree)

Total_beta_CR_pd <- Create_dissims(input_df = TPM_match_CR,
                                   input_metadata = metadata_ord_CR,
                                   input_list = shrew_list,
                                   stored_data = Total_beta_CR_pd,
                                   hillR_func = hill_phylo_parti,
                                   extra_input = T,
                                   extra_file = CR_tree)

Total_beta_pd <- rbind(Total_beta_AS_pd, Total_beta_CR_pd) %>%
  as_tibble 

Total_beta_pd <- make_overall_csv(Total_beta_pd)
write.csv(Total_beta_pd, './outputsR/total_betadiversity_pd.csv')

Overall_beta <- plot_overall_betas(Total_beta_pd)
Overall_beta
ggsave('./figures/individual_betadiversity_pd.png', width = 10, height = 7)
ggsave('./figures/individual_betadiversity_pd.pdf', width = 10, height = 7)
```

Overall functional diversity
####Functional diversity plots CR
```{r include=FALSE}
Crocidura_products <- read.csv('../distillR/outputsR/distilled_table_functions_CR.csv', sep=',', row.names = 1) %>%
  .[row.names(crocidura_tpm),]
####Check for differences in rows
setdiff(Crocidura_products$X, row.names(crocidura_tpm))

####Load the functional traits for Apodemus
#####Apodemus 
Apodemus_products <- read.csv('../distillR/outputsR/distilled_table_functions_AS.csv', sep=',', row.names = 1) %>%
  .[row.names(apodemus_tpm),]

apodemus_products <- Apodemus_products

####Check for differences in rows
setdiff(row.names(apodemus_products), row.names(apodemus_tpm))
```

Overall functional diversity plots
```{r}
Total_beta_AS_fd <- Create_dissims(input_df = TPM_match_AS,
                                   input_metadata = metadata_ord_AS,
                                   input_list = mouse_list,
                                   stored_data = Total_beta_AS_fd,
                                   hillR_func = hill_func_parti,
                                   extra_input = TRUE,
                                   extra_file = apodemus_products)

####Create the dissimilarity table
Total_beta_CR_fd <- Create_dissims(input_df = TPM_match_CR,
                                   input_metadata = metadata_ord_CR,
                                   input_list = shrew_list,
                                   stored_data = Total_beta_CR_fd,
                                   hillR_func = hill_func_parti,
                                   extra_input = TRUE,
                                   extra_file = Crocidura_products)

Total_beta_fd <- rbind(Total_beta_AS_fd, Total_beta_CR_fd) %>%
  as_tibble

Total_beta_fd <- make_overall_csv(Total_beta_fd)
write.csv(Total_beta_fd, './outputsR/total_betadiversity_fd.csv')

Overall_beta <- plot_overall_betas(Total_beta_fd)
Overall_beta
ggsave('./figures/individual_betadiversity_fd.png', width = 10, height = 7)
ggsave('./figures/individual_betadiversity_fd.pdf', width = 10, height = 7)
```

Check Alpha diversity indices across different treatments and plot the diversity metrics
```{r}
Alpha_div_td_AS <- hill_taxa(t(apodemus_tpm), q=1) %>%
  as.data.frame() %>%
  rename('diversity' = '.') %>%
  rownames_to_column('sample_id') %>%
  left_join(metadata, by = 'sample_id') %>%
  mutate(treatment = factor(treatment, levels=c('Day1','Acc','Heat','Cold','Diet')))

Alpha_div_td_CR <- hill_taxa(t(crocidura_tpm), q=1) %>%
  as.data.frame() %>%
  rename('diversity' = '.') %>%
  rownames_to_column('sample_id') %>%
  left_join(metadata, by = 'sample_id') %>%
  mutate(treatment = factor(treatment, levels=c('Day1','Acc','Heat','Cold','Diet')))

Alpha_combined <- rbind(Alpha_div_td_AS, Alpha_div_td_CR)

write.csv(Alpha_combined, './outputsR/diversity_table.csv')

alpha <- ggplot(unique(Alpha_combined), aes(x=species, y=diversity, fill=species)) +
  facet_wrap(~treatment, nrow = 1) +
  geom_boxplot(width=0.5, aes(fill=species),size=0.25) +
  geom_jitter(alpha=0.2, position = position_jitter(0.2), size=0.5) +
  scale_fill_manual(values=c('#fbc958', '#5c7f9d')) +
  theme(axis.text.x = element_text(angle=45,
                                   vjust=0.5),
        panel.background = element_blank(),
        strip.background = element_blank(),
        axis.title.x = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm"),
        legend.title = element_text(size = 12,
                                    face = 'bold'),
        legend.text = element_text(size = 10),
        axis.title.y = element_text(size = 12,
                                    face = 'bold'),
        strip.text.x = element_text(size = 8),
        legend.position = 'bottom',
        legend.key=element_rect(fill="white")) +
  labs(fill='Species') +
  ylab('Alpha diversity')
alpha
```

Alpha diversities for phylogenetic metrics
```{r}
Alpha_div_pd_AS <- hill_phylo(t(apodemus_tpm), tree = AS_tree, q=1) %>%
  as.data.frame() %>%
  rename('diversity' = '.') %>%
  rownames_to_column('sample_id') %>%
  left_join(metadata, by = 'sample_id') %>%
  mutate(treatment = factor(treatment, levels=c('Day1','Acc','Heat','Cold','Diet')))

Alpha_div_pd_CR <- hill_phylo(t(crocidura_tpm), tree = CR_tree, q=1) %>%
  as.data.frame() %>%
  rename('diversity' = '.') %>%
  rownames_to_column('sample_id') %>%
  left_join(metadata, by = 'sample_id') %>%
  mutate(treatment = factor(treatment, levels=c('Day1','Acc','Heat','Cold','Diet')))

Alpha_combined_pd <- rbind(Alpha_div_pd_AS, Alpha_div_pd_CR)

write.csv(Alpha_combined_pd, './outputsR/diversity_table_pd.csv')

alpha_pd <- ggplot(unique(Alpha_combined_pd), aes(x=species, y=diversity, fill=species)) +
  facet_wrap(~treatment, nrow = 1) +
  geom_boxplot(width=0.5, aes(fill=species), size=0.5, alpha = 0.85) +
  geom_jitter(alpha=0.2, position=position_jitter(0.2), size=0.5) +
  scale_fill_manual(values=c('#fbc958', '#5c7f9d')) +
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle=45,
                                   vjust=0.5),
        strip.background = element_blank(),
        axis.title.x = element_blank(),
        legend.position = 'none',
        plot.margin = unit(c(0,0,0,0), "cm"),
        strip.text.x = element_text(size = 8)) +
  labs(fill='Treatment') +
  ylab('')
alpha_pd
```

Alpha diversities for functional metrics
```{r}
Alpha_div_fd_AS <- hill_func(t(apodemus_tpm), apodemus_products, q=1) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column('sample_id') %>%
  left_join(metadata, by = 'sample_id') %>%
  mutate(treatment = factor(treatment, levels=c('Day1','Acc','Heat','Cold','Diet')))

Alpha_div_fd_CR <- hill_func(t(crocidura_tpm), Crocidura_products, q=1) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column('sample_id') %>%
  left_join(metadata, by = 'sample_id') %>%
  mutate(treatment = factor(treatment, levels=c('Day1','Acc','Heat','Cold','Diet')))

Alpha_combined_fd <- rbind(Alpha_div_fd_AS, Alpha_div_fd_CR)

write.csv(Alpha_combined_fd, './outputsR/diversity_table_fd.csv')

alpha_fd <- ggplot(Alpha_combined_fd, aes(x=species, y=Q, fill=species)) +
  facet_grid(~treatment, scales = 'free') +
  geom_boxplot(width=0.5, aes(fill=species), size=0.5, alpha = 0.7) +
  geom_jitter(alpha=0.2, position=position_jitter(0.2), size=0.5) +
  scale_fill_manual(values=c('#fbc958', '#5c7f9d')) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        axis.title.x = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm"),
        strip.text.x = element_text(size = 8),
        axis.text.x = element_text(angle = 45)) +
  labs(fill='Disturbance') +
  ylab('') +
  ylim(0,7.5)
alpha_fd
```

Taxonomic beta diversities
```{r}
correct_order <- c("Day_Acc_AS", "Day_Acc_CR", "Acc_Heat_AS", "Acc_Heat_CR","Heat_Cold_AS","Heat_Cold_CR","Cold_Diet_AS", "Cold_Diet_CR")

dissim_table_AS <- beta_div_func(input_df = TPM_match_AS,
                                 input_metadata = metadata_ord_AS,
                                 input_list = mouse_list,
                                 input_function = hill_taxa_parti,
                                 extra_input = F,
                                 phylo = F) %>%
  mutate(Treatment_pair = str_remove(Treatment_pair, '[0-9]$'),
         Treatment_pair = str_remove(Treatment_pair, '[0-9]'),
         Species = rep('AS'))
 

dissim_table_CR <- beta_div_func(input_df = TPM_match_CR,
                                 input_metadata = metadata_ord_CR,
                                 input_list = shrew_list,
                                 input_function = hill_taxa_parti,
                                 extra_input = F,
                                 phylo = F) %>%
  mutate(Treatment_pair = str_remove(Treatment_pair, '[0-9]$'),
         Treatment_pair = str_remove(Treatment_pair, '[0-9]'),
         Species = rep('CR'))

Beta_combined_td <- rbind(dissim_table_AS, dissim_table_CR) %>%
  mutate(treatmentWspecies = str_c(Treatment_pair, Species, sep = '_'),
         treatmentWspecies = str_replace(treatmentWspecies, 'Day_Acc1_AS', 'Day_Acc_AS')) %>%
  mutate(treatmentWspecies = str_replace(treatmentWspecies, 'Day_Acc1_CR', 'Day_Acc_CR'),
         treatmentWspecies = factor(treatmentWspecies, levels=correct_order),
         beta = 1 - local_similarity)

write.csv(Beta_combined_td, './outputsR/dissimilarity_combined_td.csv')

tax_beta <- ggplot(Beta_combined_td, aes(x=treatmentWspecies, y=beta, fill=Species)) +
  geom_boxplot(width=0.5) +
  geom_jitter(alpha=0.2, position = position_jitter(0.2), size=0.5) +
  scale_fill_manual(values=c('#fbc958', '#5c7f9d')) +
  theme(axis.text.x = element_text(angle=45,
                                   vjust=0.5,
                                   hjust=0.3),
        panel.background = element_blank(),
        axis.title.x = element_text(size=12,
                                    face='bold'),
        plot.margin = unit(c(0.75,0,0,0), "cm"),
        legend.title = element_text(size = 12,
                                    face = 'bold'),
        legend.text = element_text(size = 10),
        axis.title.y = element_text(size = 12,
                                   face = 'bold'),
        legend.position = 'bottom') +
  ylim(0, 1) +
  ylab('Beta diversity') +
  xlab('Neutral') +
  labs(fill='Species') +
  scale_x_discrete(breaks=c('Day_Acc_AS', 'Acc_Heat_AS', 'Heat_Cold_AS', 'Cold_Diet_AS'),
                   labels=c('Day1_Acc', 'Acc_Heat', 'Heat_Cold', 'Cold_Diet'))

tax_beta

```

Phylogenetic beta diversities
```{r}
dissim_table_AS_pd <- beta_div_func(input_df = TPM_match_AS,
                                 input_metadata = metadata_ord_AS,
                                 input_list = mouse_list,
                                 stored_data = dissim_table_AS,
                                 input_function = hill_phylo_parti,
                                 extra_input = T,
                                 extra_file = AS_tree,
                                 phylo = T) %>%
  mutate(Treatment_pair = str_remove(Treatment_pair, '[0-9]$'),
         Treatment_pair = str_remove(Treatment_pair, '[0-9]'),
         Species = rep('AS'))

dissim_table_CR_pd <- beta_div_func(input_df = TPM_match_CR,
                                 input_metadata = metadata_ord_CR,
                                 input_list = shrew_list,
                                 stored_data = dissim_table_CR,
                                 input_function = hill_phylo_parti,
                                 extra_input = T,
                                 extra_file = CR_tree,
                                 phylo = T) %>%
  mutate(Treatment_pair = str_remove(Treatment_pair, '[0-9]$'),
         Treatment_pair = str_remove(Treatment_pair, '[0-9]'),
         Species = rep('CR'))

Beta_combined_pd <- rbind(dissim_table_AS_pd, dissim_table_CR_pd) %>%
  mutate(treatmentWspecies = str_c(Treatment_pair, Species, sep = '_'),
         treatmentWspecies = str_replace(treatmentWspecies, 'Day_Acc1_AS', 'Day_Acc_AS')) %>%
  mutate(treatmentWspecies = str_replace(treatmentWspecies, 'Day_Acc1_CR', 'Day_Acc_CR'),
         treatmentWspecies = factor(treatmentWspecies, levels=correct_order),
         beta = 1 - local_similarity)

write.csv(Beta_combined_pd, './outputsR/Beta_combined_pd.csv')

pd_beta <- ggplot(Beta_combined_pd, aes(x=treatmentWspecies, y=beta, fill=Species)) +
  geom_boxplot(width=0.5, alpha = 0.85) +
  geom_jitter(alpha=0.2, position=position_jitter(0.2), size=0.5) +
  scale_fill_manual(values=c('#fbc958', '#5c7f9d')) +
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle=45,
                                   vjust=0.5,
                                   hjust=0.3),
        legend.position = 'none',
        axis.title.x = element_text(size=12,
                                    face = 'bold'),
        plot.margin=margin(0.75, 0, 0, 0,unit="cm"),
        axis.title.y = element_blank()) +
  ylim(0, 1) +
  labs(fill='Species') +
  xlab('Phylogenetic') +
  scale_x_discrete(breaks=c('Day_Acc_AS', 'Acc_Heat_AS', 'Heat_Cold_AS', 'Cold_Diet_AS'),
                   labels=c('Day1-Acc', 'Acc-Heat', 'Heat-Cold', 'Cold-Diet'))
pd_beta
```
Functional beta diversities
```{r}
dissim_table_AS_fd <- beta_div_func(input_df = TPM_match_AS,
                                 input_metadata = metadata_ord_AS,
                                 input_list = mouse_list,
                                 input_function = hill_func_parti,
                                 extra_input = T,
                                 extra_file = apodemus_products,
                                 phylo = F) %>%
  mutate(Treatment_pair = str_remove(Treatment_pair, '[0-9]$'),
         Treatment_pair = str_remove(Treatment_pair, '[0-9]'),
         Species = rep('AS'))

dissim_table_CR_fd <- beta_div_func(input_df = TPM_match_CR,
                                 input_metadata = metadata_ord_CR,
                                 input_list = shrew_list,
                                 input_function = hill_func_parti,
                                 extra_input = T,
                                 extra_file = Crocidura_products,
                                 phylo = F) %>%
  mutate(Treatment_pair = str_remove(Treatment_pair, '[0-9]$'),
         Treatment_pair = str_remove(Treatment_pair, '[0-9]'),
         Species = rep('CR'))

Beta_combined_fd <- rbind(dissim_table_AS_fd, dissim_table_CR_fd) %>%
  mutate(treatmentWspecies = str_c(Treatment_pair, Species, sep = '_'),
         treatmentWspecies = str_replace(treatmentWspecies, 'Day_Acc1_AS', 'Day_Acc_AS')) %>%
  mutate(treatmentWspecies = str_replace(treatmentWspecies, 'Day_Acc1_CR', 'Day_Acc_CR'),
         treatmentWspecies = factor(treatmentWspecies, levels=correct_order),
         beta = 1 - local_similarity)

write.csv(Beta_combined_fd, './outputsR/dissimilarity_combined_fd.csv')

fd_beta <- ggplot(Beta_combined_fd, aes(x=treatmentWspecies, y=beta, fill=Species)) +
  geom_boxplot(width=0.5, alpha =0.7) +
  geom_jitter(alpha=0.2, position=position_jitter(0.2), size=0.5) +
  scale_fill_manual(values=c('#fbc958', '#5c7f9d')) +
  theme(panel.background = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_text(angle = 45,
                                   vjust=0.5,
                                   hjust=0.3),
        axis.title.x = element_text(size = 12,
                                    face = 'bold'),
        legend.position = 'none',
        plot.margin = unit(c(0.75,0,0,0), "cm"),
        axis.title.y = element_blank()) +
  labs(fill='Species') +
  xlab('Functional') +
  ylim(c(0,1)) +
  scale_x_discrete(breaks=c('Day_Acc_AS', 'Acc_Heat_AS', 'Heat_Cold_AS', 'Cold_Diet_AS'),
                   labels=c('Day1-Acc', 'Acc-Heat', 'Heat-Cold', 'Cold-Diet'))

fd_beta

```

```{r}
rm_legend <- function(p){p + theme(legend.position = "none")}

alpha_beta_plots <- ggarrange(rm_legend(alpha), rm_legend(alpha_pd), rm_legend(alpha_fd), tax_beta, rm_legend(pd_beta),  rm_legend(fd_beta), 
          labels=c('A', 'B', 'C', 'D', 'E', 'F'),
          ncol=3, nrow=2,
          common.legend = T,
          legend = 'bottom')
alpha_beta_plots

ggsave('./figures/alpha_beta_plots.pdf', alpha_beta_plots, height = 7, width = 10, dpi = 1000)
ggsave('./figures/alpha_beta_plots.png', alpha_beta_plots, height = 7, width = 10, dpi = 1000)
ggsave('./figures/alpha_beta_plots.tiff', compression = 'lzw')

```

```{r}
sessionInfo()
```
