---
title: "linear_models on alpha hill numbers"
author: "Adam_Koziol"
date: "4/21/2022"
output:
  pdf_document: default
  html_document: default
---
setting up the environment    

```{r include=FALSE}
setwd('~/Documents/GitHub/Plasticity_Exp_Repo/linear_models/')
library(tidyverse)
library(nlme)
library(broom)
library(ggplotify)
library(multcomp)
library(tinytex)
library(lme4)
library(jtools)
library(sjPlot)
library(effects)
```
Loading the data
```{r}
####For some reason the dataframe has duplicates - remove these
alpha_diversity_tax <- read.csv('../hill_numbers/outputsR/diversity_table.csv', header=T, row.names=1) %>%
  unique(.) %>%
  as.tibble
#####Combined dataframe containing functional diversities
alpha_diversity_fd <- read.csv('../hill_numbers/outputsR/diversity_table_fd.csv', header=T, row.names = 1)
####Phylogenetic diversities
alpha_diversity_pd <- read.csv('../hill_numbers/outputsR/diversity_table_pd.csv', header=T, row.names = 1)
#####Metadata
metadata <- read.csv('../data/metadata.csv')
```
Combining the metadata with the dataframes
```{r}
alph_div_tax_as <- alpha_diversity_tax %>%
  filter(species == 'AS') %>% 
  dplyr::select(c(diversity, sample_id)) %>%
  left_join(metadata, by='sample_id') %>%
  mutate(days=case_when(treatment == 'Day1' ~ 0,
                        treatment == 'Acc' ~ 7,
                        treatment == 'Heat' ~ 19,
                        treatment == 'Cold' ~ 33,
                        treatment == 'Diet' ~ 45)) %>%
  mutate(treatment = factor(treatment, levels = c('Day1', 'Acc', 'Heat', 'Cold', 'Diet')))

alph_div_tax_cr <- alpha_diversity_tax %>%
  filter(species == 'CR') %>%
  dplyr::select(c(diversity, sample_id)) %>%
  left_join(metadata, by='sample_id') %>%
  mutate(days=case_when(treatment == 'Day1' ~ 0,
                        treatment == 'Acc' ~ 7,
                        treatment == 'Heat' ~ 19,
                        treatment == 'Cold' ~ 33,
                        treatment == 'Diet' ~ 45),
         cage=case_when(cage == 1 ~ 6,
                        cage == 2 ~ 7,
                        cage == 3 ~ 8,
                        cage == 4 ~ 9,
                        cage == 5 ~ 10,
                        cage == 6 ~ 11)) %>%
  mutate_at(2:8, as.factor) %>%
  mutate(treatment = factor(treatment, levels = c('Day1', 'Acc', 'Heat', 'Cold', 'Diet')))

alph_div_fd_as <- alpha_diversity_fd %>%
  filter(species == 'AS') %>%
  dplyr::select(c(Q, sample_id)) %>%
  left_join(metadata, by='sample_id') %>%
  filter(treatment != 'FAST') %>%
  mutate(days=case_when(treatment == 'Day1' ~ 0,
                        treatment == 'Acc' ~ 7,
                        treatment == 'Heat' ~ 19,
                        treatment == 'Cold' ~ 33,
                        treatment == 'Diet' ~ 45)) %>%
  mutate_at(2:8, as.factor) %>%
  mutate(treatment = factor(treatment, levels = c('Day1', 'Acc', 'Heat', 'Cold', 'Diet')))

alph_div_fd_cr <- alpha_diversity_fd %>%
  filter(species == 'CR') %>%
  dplyr::select(c(Q, sample_id)) %>%
  left_join(metadata, by='sample_id') %>%
  mutate(days=case_when(treatment == 'Day1' ~ 0,
                        treatment == 'Acc' ~ 7,
                        treatment == 'Heat' ~ 19,
                        treatment == 'Cold' ~ 33,
                        treatment == 'Diet' ~ 45),
         cage=case_when(cage == 1 ~ 6,
                        cage == 2 ~ 7,
                        cage == 3 ~ 8,
                        cage == 4 ~ 9,
                        cage == 5 ~ 10,
                        cage == 6 ~ 11)) %>%
  mutate_at(2:8, as.factor) %>%
  mutate(treatment = factor(treatment, levels = c('Day1', 'Acc', 'Heat', 'Cold', 'Diet')))

alph_div_pd_as <- alpha_diversity_pd %>%
  filter(species == 'AS') %>%
  dplyr::select(c(diversity, sample_id)) %>%
  left_join(metadata, by='sample_id') %>%
  filter(treatment != 'FAST') %>%
  mutate(days=case_when(treatment == 'Day1' ~ 0,
                        treatment == 'Acc' ~ 7,
                        treatment == 'Heat' ~ 19,
                        treatment == 'Cold' ~ 33,
                        treatment == 'Diet' ~ 45)) %>%
  mutate_at(2:8, as.factor) %>%
  mutate(treatment = factor(treatment, levels = c('Day1', 'Acc', 'Heat', 'Cold', 'Diet')))

alph_div_pd_cr <- alpha_diversity_pd %>%
  filter(species == 'CR') %>%
  dplyr::select(c(diversity, sample_id)) %>%
  left_join(metadata, by='sample_id') %>%
  mutate(days=case_when(treatment == 'Day1' ~ 0,
                        treatment == 'Acc' ~ 7,
                        treatment == 'Heat' ~ 19,
                        treatment == 'Cold' ~ 33,
                        treatment == 'Diet' ~ 45),
         cage=case_when(cage == 1 ~ 6,
                        cage == 2 ~ 7,
                        cage == 3 ~ 8,
                        cage == 4 ~ 9,
                        cage == 5 ~ 10,
                        cage == 6 ~ 11)) %>%
  mutate_at(2:8, as.factor) %>%
  mutate(treatment = factor(treatment, levels = c('Day1', 'Acc', 'Heat', 'Cold', 'Diet')))
```
Sanity check of each factor
```{r}
sapply(alph_div_tax_as, mode)
contr <- rbind("Day1-Acc" = c(1, -1, 0, 0, 0),
               "Acc-Heat" = c(0, 1, -1, 0, 0),
               "Heat-Cold" = c(0, 0, 1,-1, 0),
               "Cold-Diet" = c(0, 0, 0, 1,-1))
contr_cr <- rbind("Day1-Acc" = c(1, -1, 0, 0, 0),
               "Acc-Heat" = c(0, 1, -1, 0, 0),
               "Heat-Cold" = c(0, 0, 1,-1, 0),
               "Cold-Diet" = c(0, 0, 0, 1,-1))
```

```{r}
hist(alph_div_tax_as$diversity)
```

treatment_wise_boxplots from the taxonomic diversity Hill numbers 
```{r}
boxplot(diversity~treatment*cage,data = alph_div_tax_as)
```
Assessment of log transformation it looks tighter, however the non-transformed data still looks ok to me so i will continue with this
```{r}
boxplot(log(diversity)~treatment*cage,data = alph_div_tax_as)
```
Create the models for comparison - including the random effects results in the lowest AIC
```{r}
###Try to include the other factors 
model_as.a <- lme(diversity~treatment,
                  random = ~1|cage/individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_tax_as)
model_as.b <- lme(diversity~treatment,
                  random = ~1|cage,
                  data=alph_div_tax_as)
model_as.c <- lme(diversity~treatment,
                  random = ~1|individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_tax_as)
#model_as.d <- lme(diversity~treatment,
                  #random = ~1|cage,
                  #correlation = corCAR1(form = ~ days),
                  #data=alph_div_tax_as)
model_as.e <- lm(diversity~treatment,
                  data=alph_div_tax_as)

anova(model_as.a,model_as.b,model_as.c,model_as.e)
AIC(model_as.a) #931.8822
AIC(model_as.b) #933.4933
AIC(model_as.c) #930.3255
AIC(model_as.e) #960.1475
```
Residuals looks like they are randomly distributed around 0
```{r}
model_as1=model_as.a
E <- resid(model_as1, type = "normalized")
Fit <- fitted(model_as1)
op <- par(mfrow = c(1, 2))
plot(x = Fit, y = E,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals versus fitted values")
hist(E, nclass = 15)
```


```{r}
par(op)
bp=boxplot(E ~ treatment,
           data = alph_div_tax_as)
```
```{r}
summary(model_as.a)
```

We have a slightly significant treatment effect
```{r}
anova(model_as.a)
```
```{r}
plot(allEffects(model_as.a))
```

Pairwise comparison of treatments to assess where the significant effects are
```{r}
summary(glht(model_as.a, linfct = mcp(treatment=contr),
                   test = adjusted("holm")))
```

```{r}
#plot_model(model_as1, type='Day1d', terms = c('treatment'))
```

################Start if the Crocidura taxonomic linear models########################
```{r}
boxplot(diversity~treatment*cage,data = alph_div_tax_cr)
```

```{r}
boxplot(log(diversity)~treatment*cage,data = alph_div_tax_cr)
```

```{r}
hist(alph_div_tax_cr$diversity)
```
The lowest AIC is the gls without random effects - however i consider the random effects to be an essential study design parameter - i'll continue with the random effects
included
```{r}
#model_cr.a <- lme(diversity~treatment,
                  #random = ~1|cage/individual_id,
                  #correlation = corCAR1(form = ~ as.numeric(days)),
                  #data=alph_div_tax_cr)
model_cr.b <- lme(diversity~treatment,
                  random = ~1|cage,
                  data=alph_div_tax_cr)
model_cr.c <- lme(diversity~treatment,
                  random = ~1|individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_tax_cr)
model_cr.e <- lm(diversity~treatment,
                  data=alph_div_tax_cr)

#AIC(model_cr.a) #657.3954
AIC(model_cr.b) #654.6046
AIC(model_cr.c) #655.3954
#AIC(model_cr.d) #NA
AIC(model_cr.e) #658.816
```




```{r}
model_cr1=model_cr.b
E <- resid(model_cr1, type = "normalized")
Fit <- fitted(model_cr1)
op <- par(mfrow = c(1, 2))
plot(x = Fit, y = E,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals versus fitted values")
hist(E, nclass = 15)
```
```{r}
par(op)
bp=boxplot(E ~ treatment,
           data = alph_div_tax_cr)
```

Again we have a slightly significance treatment effect on taxonomy
```{r}
anova(model_cr1)
```
Pairwise comparison of treatments
```{r}
summary(glht(model_cr1, linfct = mcp(treatment=contr_cr),
                   test = adjusted("holm")))
```
################Start of functional alpha diversity Apodemus#################
```{r}
boxplot(Q~treatment*cage,data = alph_div_fd_as)
```

```{r}
boxplot(log(Q)~treatment*cage,data = alph_div_fd_as)
```
```{r}
hist(alph_div_fd_as$Q)
```



```{r}
model_as.a_fd <- lme(Q~treatment,
                  random = ~1|cage/individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_fd_as)
model_as.b_fd <- lme(Q~treatment,
                  random = ~1|individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_fd_as)
model_as.c_fd <- lme(Q~treatment,
                  random = ~1|cage,
                  data=alph_div_fd_as)
model_as.d_fd <- lme(Q~treatment,
                  random = ~1|cage/individual_id,
                  data=alph_div_fd_as)
model_as.e_fd <- lm(Q~treatment,
                  data=alph_div_fd_as)
AIC(model_as.a_fd) #292.0188
AIC(model_as.b_fd) #290.6965
AIC(model_as.c_fd) #292.2867
AIC(model_as.d_fd) #292.3475
AIC(model_as.e_fd) #286.632
```


```{r}
model_as1_fd=model_as.a_fd
E <- resid(model_as1_fd, type = "normalized")
Fit <- fitted(model_as.a_fd)
op <- par(mfrow = c(1, 2))
plot(x = Fit, y = E,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals versus fitted values")
hist(E, nclass = 15)
par(op)
bp=boxplot(E ~ treatment,
           data = alph_div_fd_as)
```
We observe no significant difference in the effect of treatment on functional alpha diversity
```{r}
anova(model_as1_fd)
```
This is useless - ignore it for now but just wanted to sanity check
```{r}
summary(glht(model_as1_fd, linfct = mcp(treatment=contr),
                   test = adjusted("holm")))
```
####################Start of the Crocidura functional alpha diversity
```{r}
boxplot(Q~treatment*cage,data = alph_div_fd_cr)
```

```{r}
boxplot(log(Q)~treatment*cage,data = alph_div_fd_cr)
```
```{r}
hist(alph_div_fd_cr$Q)
```


```{r}
model_cr.a_fd <- lme(Q~treatment,
                  random = ~1|cage/individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_fd_cr)
model_cr.b_fd <- lme(Q~treatment,
                  random = ~1|individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_fd_cr)
model_cr.c_fd <- lme(Q~treatment,
                  random = ~1|cage,
                  data=alph_div_fd_cr)
model_cr.d_fd <- lme(Q~treatment,
                  random = ~1|individual_id,
                  data=alph_div_fd_cr)
model_cr.e_fd <- lm(Q~treatment,
                  data=alph_div_fd_cr)


AIC(model_cr.a_fd) #377.5765
AIC(model_cr.b_fd) #376.0824
AIC(model_cr.c_fd) #379.5765
AIC(model_cr.d_fd) #374.0913
AIC(model_cr.e_fd) #370.5743
```

```{r}
model_cr1_fd=model_cr.a_fd
E <- resid(model_cr1_fd, type = "normalized")
Fit <- fitted(model_cr1_fd)
op <- par(mfrow = c(1, 2))
plot(x = Fit, y = E,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals versus fitted values")
hist(E, nclass = 15)
par(op)
bp=boxplot(E ~ treatment,
           data = alph_div_tax_cr)
```
We have a highly significant effect of treatment on functional alpha diversity
```{r}
anova(model_cr1_fd)
```

```{r}
summary(glht(model_cr1_fd, linfct = mcp(treatment=contr_cr),
                   test = adjusted("holm")))
```
```{r}
plot(allEffects(model_cr1_fd))
```


########Phylogenetic linear models
```{r}
hist(alph_div_pd_as$diversity)
```

treatment_wise_boxplots from the taxonomic diversity Hill numbers 
```{r}
boxplot(diversity~treatment*cage,data = alph_div_pd_as)
```
Assessment of log transformation it looks tighter, however the non-transformed data still looks ok to me so i will continue with this
```{r}
boxplot(log(diversity)~treatment*cage,data = alph_div_pd_as)
```
Create the models for comparison - including the random effects results in the lowest AIC
```{r}
###Try to include the other factors 
model_as.a.pd <- lme(diversity~treatment,
                  random = ~1|cage/individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_pd_as)
model_as.b.pd  <- lme(diversity~treatment,
                  random = ~1|individual_id,
                  data=alph_div_pd_as)
model_as.c.pd  <- lme(diversity~treatment,
                  random = ~1|individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_pd_as)
model_as.d.pd  <- lme(diversity~treatment,
                  random = ~1|cage,
                  data=alph_div_pd_as)
model_as.e.pd  <- lm(diversity~treatment,
                  data=alph_div_pd_as)

anova(model_as.a.pd,model_as.b.pd,model_as.c.pd,model_as.e.pd)
AIC(model_as.a.pd ) #525.004
AIC(model_as.b.pd ) #529.4421
AIC(model_as.c.pd ) #529.9477
AIC(model_as.d.pd ) #523.7998
AIC(model_as.e.pd ) #546.8065
```
Residuals looks like they are randomly distributed around 0
```{r}
model_as1_pd=model_as.a.pd
E <- resid(model_as1_pd, type = "normalized")
Fit <- fitted(model_as1_pd)
op <- par(mfrow = c(1, 2))
plot(x = Fit, y = E,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals versus fitted values")
hist(E, nclass = 15)
```


```{r}
par(op)
bp=boxplot(E ~ treatment,
           data = alph_div_pd_as)
```
```{r}
summary(model_as1_pd)
#####There is an interesting sex effect on Phylogenetic diversity sexMALE	-5.406496	1.414101	3	-3.823273	0.0315
```

We have a slightly significant treatment effect
```{r}
anova(model_as1_pd)

#There is a significant treatment and sex effect 
#Treatment = 0.0374
#Sex = 0.0105
```

```{r}
plot(allEffects(model_as1_pd))
```


######Phylogenetic on the Crocidura phylogenetic diversity

```{r}
hist(alph_div_pd_cr$diversity)
```

treatment_wise_boxplots from the taxonomic diversity Hill numbers 
```{r}
boxplot(diversity~treatment*cage,data = alph_div_pd_cr)
```
Assessment of log transformation it looks tighter, however the non-transformed data still looks ok to me so i will continue with this
```{r}
boxplot(log(diversity)~treatment*cage,data = alph_div_pd_cr)
```
Create the models for comparison - including the random effects results in the lowest AIC
```{r}
###Try to include the other factors 
model_cr.a.pd <- lme(diversity~treatment,
                  random = ~1|cage/individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_pd_cr)
model_cr.b.pd  <- lme(diversity~treatment,
                  random = ~1|individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_pd_cr)
model_cr.c.pd  <- lme(diversity~treatment,
                  random = ~1|cage,
                  data=alph_div_pd_cr)
model_cr.d.pd  <- lme(diversity~treatment,
                  random = ~1|individual_id,
                  data=alph_div_pd_cr)
model_cr.e.pd <-  lm(diversity~treatment,
                     data=alph_div_pd_cr)


anova(model_cr.a.pd,model_cr.b.pd,model_cr.c.pd,model_cr.e.pd)
AIC(model_cr.a.pd ) #383.197
AIC(model_cr.b.pd ) #381.197
AIC(model_cr.c.pd ) #379.439
AIC(model_cr.d.pd ) #379.439
AIC(model_cr.e.pd ) #377.9547
```
Residuals looks like they are randomly distributed around 0
```{r}
model_cr1_pd=model_cr.a.pd
E <- resid(model_cr1_pd, type = "normalized")
Fit <- fitted(model_cr1_pd)
op <- par(mfrow = c(1, 2))
plot(x = Fit, y = E,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals versus fitted values")
hist(E, nclass = 15)
```
```{r}
plot(model_cr1_pd)
```


```{r}
par(op)
bp=boxplot(E ~ treatment,
           data = alph_div_pd_cr)
```
```{r}
summary(model_cr1_pd)
#####There is an interesting sex effect on Phylogenetic diversity sexMALE	-5.406496	1.414101	3	-3.823273	0.0315
```


We have a slightly significant treatment effect
```{r}
anova(model_cr1_pd)
```
```{r}
plot(allEffects(model_cr1_pd))
```
Species-species alpha diversity linear model
```{r}
alph_div_combined <- rbind(alph_div_tax_as, alph_div_tax_cr)

hist(alph_div_combined$diversity)
```
```{r}
boxplot(diversity~treatment*cage,data = alph_div_combined)
```
```{r}
boxplot(log10(diversity)~treatment*cage,data = alph_div_combined)
```

```{r}
###Try to include the other factors 
model_div_combined.a <- lme(log(diversity)~species,
                  random = ~1|cage/individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_combined)
model_div_combined.b  <- lme(diversity~species,
                  random = ~1|individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_combined)
model_div_combined.c  <- lme(diversity~species,
                  random = ~1|cage,
                  data=alph_div_combined)
model_div_combined.d  <- lme(diversity~species,
                  random = ~1|individual_id,
                  data=alph_div_combined)
model_div_combined.e <-  lm(diversity~species,
                     data=alph_div_combined)


AIC(model_div_combined.a) #1868.937
AIC(model_div_combined.b) #1868.075
AIC(model_div_combined.c) #1875.097
AIC(model_div_combined.d) #1872.694
AIC(model_div_combined.e) #1885.803
```
```{r}
model_alph_combined=model_div_combined.a
E <- resid(model_alph_combined, type = "normalized")
Fit <- fitted(model_alph_combined)
op <- par(mfrow = c(1, 2))
plot(x = Fit, y = E,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals versus fitted values")
hist(E, nclass = 15)
```
```{r}
plot(model_alph_combined)
```
```{r}
#par(op)
#bp=boxplot(E ~ species,
#           data = model_div_combined.a)
```
```{r}
summary(model_alph_combined)
```
```{r}
anova(model_alph_combined)
```

```{r}
plot(allEffects(model_alph_combined))
```

```{r}
###Try to include the other factors 
alph_div_fd_combined <- rbind(alph_div_fd_as, alph_div_fd_cr)
hist(alph_div_fd_combined$Q)
boxplot(Q~treatment*cage,data = alph_div_fd_combined)
boxplot(log10(Q)~treatment*cage,data = alph_div_fd_combined)
model_div_fd_combined.a <- lme(Q~species,
                  random = ~1|cage/individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_fd_combined)
model_div_fd_combined.b  <- lme(Q~species,
                  random = ~1|individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_fd_combined)
model_div_fd_combined.c  <- lme(Q~species,
                  random = ~1|cage,
                  data=alph_div_fd_combined)
model_div_fd_combined.d  <- lme(Q~species,
                  random = ~1|individual_id,
                  data=alph_div_fd_combined)
model_div_fd_combined.e <-  lm(Q~species,
                     data=alph_div_fd_combined)


AIC(model_div_fd_combined.a) #940.832
AIC(model_div_fd_combined.b) #939.4217
AIC(model_div_fd_combined.c) #941.6157
AIC(model_div_fd_combined.d) #937.4217
AIC(model_div_fd_combined.e) #939.4188
```
```{r}
model_alph_fd_combined=model_div_fd_combined.a
E <- resid(model_alph_fd_combined, type = "normalized")
Fit <- fitted(model_alph_fd_combined)
op <- par(mfrow = c(1, 2))
plot(x = Fit, y = E,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals versus fitted values")
hist(E, nclass = 15)
```
```{r}
plot(model_alph_fd_combined)
```
```{r}
#par(op)
#bp=boxplot(E ~ species,
#           data = model_div_combined.a)
```
```{r}
summary(model_alph_fd_combined)
```
```{r}
anova(model_alph_fd_combined)
```

```{r}
plot(allEffects(model_alph_fd_combined))
```

```{r}
###Try to include the other factors 
alph_div_pd_combined <- rbind(alph_div_pd_as, alph_div_pd_cr)
hist(alph_div_pd_combined$diversity)
boxplot(diversity~treatment*cage,data = alph_div_pd_combined)
boxplot(log10(diversity)~treatment*cage,data = alph_div_pd_combined)
model_div_pd_combined.a <- lme(diversity~species,
                  random = ~1|cage/individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_pd_combined)
model_div_pd_combined.b  <- lme(diversity~species,
                  random = ~1|individual_id,
                  correlation = corCAR1(form = ~ as.numeric(days)),
                  data=alph_div_pd_combined)
model_div_pd_combined.c  <- lme(diversity~species,
                  random = ~1|cage,
                  data=alph_div_pd_combined)
model_div_pd_combined.d  <- lme(diversity~species,
                  random = ~1|individual_id,
                  data=alph_div_pd_combined)
model_div_pd_combined.e <-  lm(diversity~species,
                     data=alph_div_pd_combined)


AIC(model_div_pd_combined.a) #940.832
AIC(model_div_pd_combined.b) #939.4217
AIC(model_div_pd_combined.c) #941.6157
AIC(model_div_pd_combined.d) #937.4217
AIC(model_div_pd_combined.e) #939.4188
```
```{r}
model_alph_pd_combined=model_div_pd_combined.a
E <- resid(model_alph_pd_combined, type = "normalized")
Fit <- fitted(model_alph_pd_combined)
op <- par(mfrow = c(1, 2))
plot(x = Fit, y = E,
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals versus fitted values")
hist(E, nclass = 15)
```
```{r}
plot(model_alph_pd_combined)
```
```{r}
#par(op)
#bp=boxplot(E ~ species,
#           data = model_div_combined.a)
```
```{r}
summary(model_alph_pd_combined)
```
```{r}
anova(model_alph_pd_combined)
```

```{r}
plot(allEffects(model_alph_pd_combined))
```

```{r}
sessionInfo()
```

